A Journey from Wild to Textbook Data to Reproducibly Refresh the Wages Data from the National Longitudinal Survey of Youth Database

### 2022-02-14 paper/paper.Rmd
inconsistency check required
an attempt in fixing the outlier section

For `st_work` variable,  we only rename the column names from `new_data_qnames`. This variable is then used to calculate the next variable, `yr_workforce` for each survey round, which is the year of survey (`year`) minus the year of individual start working (`st_work`). Finally, `exp` variable is derived from the number of weeks worked since the last interview indicated as variable started with `WKSWK` in `new_data_qnames`. To obtain the work experience since 1979, we calculate the cumulative value. As the measurement unit is in week, we convert this to year.

@@

<!--AE comment: Section 3.2.1 begins “As part of the IDA, which is the model formulation, we build a robust linear regression model to treat the extreme values in the data.” This could use clarification. In 3.2, you said that part of IDA is “ model formulation without any formal statistical inference.” To me, that sounds like using modeling as a descriptive technique, and it’s something I would consider to be more a part of EDA. But then in 3.2.1, while it is perhaps not using modeling in an inferential way, the model is being used for more than just description. I might rephrase the first sentence as follows: “In order to treat the extreme values in the data, we built a robust linear regression model.” -->
As part of the IDA, <!--AE comment: Awkward sentence. "which is the model formulation"?--> which is the model formulation, we build a robust linear regression model to treat the extreme values in the data. Robust linear regression yields an estimation robust to the influence of noise or contamination [@KollerManuel2016rARP]. It also aims to detect the contamination by weighting each observation based on how "well-behaved" they are, known as robustness weight. An observation with a lower robustness weight would be suggested as an outlier [@KollerManuel2016rARP].

Since we work with longitudinal data, we build the model for each ID instead of the overall data. The robust mixed model could be the best model to be employed in this case. However, this method is too computationally and memory expensive, especially for a large data set, like the NLSY79 data. <!--I would pull the tidy and purrr citations to the end of the sentence to make it read more smoothly-->Therefore, the model for each ID is built utilizing the `nest` and `map` function from `tidyr` [@tidyr] and `purrr` [@purrr], respectively. The full code for this is shown at \url{`r yowie_pkgdown`/articles/input-to-valid-data.html} but also described in detail next.

We build the model using the `rlm` function from `MASS` package [@mass]. We set the `mean_hourly_wage` and `year` as the dependent and predictor, respectively. Furthermore, we use M-Estimation with Huber weighting, where the observation with a slight residual gets a weight of 1, while the larger the residual, the smaller the weight (less than 1) [@rlm]. <!--AE comment: “However, the challenging part of detecting the anomaly using the robustness weight is determining the weight threshold in which the observations are considered outliers.” I don’t think this sentence needs a “However” and probably should say “past which” rather than “in which”-->However, the challenging part of detecting the anomaly using the robustness weight is determining the weight threshold in which the observations are considered outliers. Moreover, it should be noted that not all the outliers are due to an error. Instead, it might be that one had reasonably increasing or decreasing wages in a particular period.

@@

To minimize the risk of mistakenly identifying an outlier as an "erroneous outlier", we simulate some thresholds and study how they affect the data. We find that 0.12 is the most reasonable value to be the threshold to minimize that drawback's risk because it still captures the sensible spikes in the data. In other words, we keep maintaining the natural variability of the wages while minimizing anomalies because of the error in the data recording. After deciding the threshold, we impute the observations whose weights are less than 0.12 with the models' predicted value. We then flag those observations in a new variable called `is_pred`.

>>>>

For `st_work` variable,  we only rename the column names from `new_data_qnames`. This variable is then used to calculate the next variable, `yr_workforce` for each survey round, which is the year of survey (`year`) minus the year of individual start working (`st_work`). Finally, `exp` variable is derived from the number of weeks worked since the last interview indicated as variable started with `WKSWK` in `new_data_qnames`. To obtain the work experience since 1979, we calculate the cumulative value. As the measurement unit is in week, we convert this to year.  [ET: inconsistency in `st_work` and `yr_workforce`?]

@@

<!--AE comment: Section 3.2.1 begins “As part of the IDA, which is the model formulation, we build a robust linear regression model to treat the extreme values in the data.” This could use clarification. In 3.2, you said that part of IDA is “ model formulation without any formal statistical inference.” To me, that sounds like using modeling as a descriptive technique, and it’s something I would consider to be more a part of EDA. But then in 3.2.1, while it is perhaps not using modeling in an inferential way, the model is being used for more than just description. I might rephrase the first sentence as follows: “In order to treat the extreme values in the data, we built a robust linear regression model.” -->
<!--AE comment: Awkward sentence. "which is the model formulation"?--> To treat the extreme values in the data, we build a robust linear regression model where we use the robustness weight to determine if a value should be replaced with the fitted value from the model. We build the model for each ID utilizing the `nest` and `map` function from `tidyr` [@tidyr] and `purrr` [@purrr], respectively. This fits a model separately for each individual, and a robust linear mixed model could be used as an alternative approach. We tried fitting a robust linear mixed model using `robustlmm` [@KollerManuel2016rARP], however perhaps because the data is large and imbalanced, the model was not converging. <!--I would pull the tidy and purrr citations to the end of the sentence to make it read more smoothly-->The full code for this is shown at \url{`r yowie_pkgdown`/articles/input-to-valid-data.html} but also described in detail next.

We build the model using the `rlm` function from `MASS` package [@mass]. We set the `mean_hourly_wage` and `year` as the dependent and predictor, respectively. Furthermore, we use M-Estimation with Huber weighting, where the observation with a slight residual gets a weight of 1, while the larger the residual, the smaller the weight (less than 1) [@rlm]. <!--AE comment: “However, the challenging part of detecting the anomaly using the robustness weight is determining the weight threshold in which the observations are considered outliers.” I don’t think this sentence needs a “However” and probably should say “past which” rather than “in which”-->The challenging part of detecting the anomaly using the robustness weight is determining the weight threshold in which the observations are considered outliers. Moreover, it should be noted that not all the outliers are due to an error. Instead, it might be that one had reasonably increasing or decreasing wages in a particular period.

@@

To minimize the risk of mistakenly identifying an outlier as an "erroneous outlier", we simulate some thresholds and study how they affect the data by eye balling the results. Based on this eye balling, we find that 0.12 is the most reasonable value to use as the threshold to minimize the risk of overly smoothing the wage by retaining sensible spikes in the data. In other words, we keep maintaining the natural variability of the wages while minimizing anomalies because of the error in the data recording. After deciding the threshold, we impute the observations whose weights are less than 0.12 with the models' predicted value. We then flag those observations in a new variable called `is_pred`.
